The complex network of trophic interactions in a subAntarctic oceanic Marine Protected Area

### 2023-04-26 MS_ToSubmit/MS_Burdwood_foodweb.tex
Updates co-authors

\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={The complex network of trophic interactions in a subAntarctic Marine Protected Area},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}

@@

  \title{The complex network of trophic interactions in a subAntarctic
Marine Protected Area}
    \author[]{Tomás I. Marina\textsuperscript{a}%
  %
  }
   \ead{tomasimarina@gmail.com} 
    \author[]{Irene R. Schloss\textsuperscript{a,b,c}%
  %
  }

@@

  \begin{abstract}
  Globally, the total area of the world ocean designated under marine
  protection has increased in the recent decades. The majority of these
  Marine Protected Areas (MPAs) focus on the presence of particularly
  vulnerable, keystone, or charismatic species, the presence of large
  numbers of endemic species, and/or the presence of high biodiversity
  across taxonomic levels. In the sub-Antarctic region, the level of
  ocean protection is mainly associated to oceanic islands, except for
  the MPAs Namuncurá - Banco Burdwood I and II (MPA N-BB, 53º--55ºS and
  56º--62ºW), which are associated to a submarine plateau and its
  southern adjacent deep slope region, respectively. Here we present the
  first detailed analysis of the network of predator-prey interactions
  (food web) for the MPA N-BB ecosystem. We applied a network approach
  with the objective of characterizing the food web in terms of
  complexity and structure, and identifying the species' role in such
  framework. In terms of complexity, the MPA N-BB food web consisted of
  1778 predator-prey interactions and 379 species, with a link density
  of 4.69 and a connectance of 0.01. In terms of structure, almost half
  of the consumers were omnivores (0.48), and the network displayed a
  small-world pattern. All this suggest that the ecosystem might be
  fragile to external perturbations targeting highly connected species,
  although structural properties might provide resilience and resistance
  with the final outcome of a rearrange structure maintaining their
  functions. On the other hand, we identified several species as
  important in terms of different aspects of trophic structure and
  functioning, and response to perturbations. We suggest that generalist
  species, mostly fishes, play a crucial role in the ecosystem's
  bentho-pelagic coupling process. In addition, we propose that other
  species rather than the longtail southern cod \emph{Patagonotothen
  ramsayi} and the Fueguian sprat \emph{Sprattus fuegensis} should be
  considered as relevant energy transfers for the ecosystem. Finally, we
  argue that it is the diversity of species, representing the benthic
  and pelagic habitats, what maintains the connectivity of the food web
  against perturbations, therefore contributing to the structure and
  stability of the ecosystem.
  \end{abstract}
    \begin{keyword}
    Food web \sep Complexity \sep Structure \sep Marine Protected
Area \sep 
    Southwest Atlantic
  \end{keyword}

@@

The evidence of benefits provided by Marine Protected Areas (MPAs) as
well as the urgent need for ocean protection have driven an
unprecedented increase in the number of MPAs worldwide in recent years
(Roberts et al., 2017; Sala et al., 2018). Globally, the total area of
the world ocean designated under marine protection adds up to nearly
29,600,000 km\textsuperscript{2}, distributed across nearly 18,444 MPAs
and covering 8.16\% of the ocean's surface (IUCN \& UNEP-WCMC, 2023),
and therefore approaching the 10\% goal of the Convention of Biological
Diversity (CBD, 2004). Despite this progress, recent reports have shown
that actual protection has been overestimated because it includes areas
that are not yet effectively protected (only declared) as well as areas
that allow significant extractive activities (Sala et al., 2018).

In the sub-Antarctic region, the level of ocean protection is mainly
associated to oceanic islands, such as the South Georgia and South
Sandwich, Bouvet, Prince Edward, and Macquarie islands (IUCN \&
UNEP-WCMC, 2023). Interestingly, the case of the MPAs Namuncurá - Banco
Burdwood I and II (MPA N-BB), which is the focus of this work, is unique
since these MPAs are associated to a submarine plateau and its southern
adjacent deep slope region, respectively (Falabella, 2017). Added to
this, such MPAs are part of a network of protected areas in the
sub-Antarctic (jointly with MPA Yaganes), that aims to protect and
strengthen the functioning of this southern region to maintain global
ocean health.

Many of these MPAs focus on the presence of particularly vulnerable,
keystone, or charismatic species, the presence of large numbers (or
proportions) of endemic species, and/or the presence of high
biodiversity across taxonomic levels (Hogg et al., 2016). Indeed, the
MPA N-BB was created to protect a sensitive and biodiverse benthic
habitat that was barely known (Falabella, 2017; L. Schejter et al.,
2016). The benthic community is featured by a high biomass of vulnerable
and fragile species (mainly Poriferaand Cnidaria) that meet the
characteristics of vulnerable marine ecosystems (Schejter \& Albano,
2021), here defined as sites that present taxa with a minimum biomass of
10 kg per 1200 m\textsuperscript{2} (Commission for the Conservation of
Antarctic Marine Living Resources, 2009). Also, it provides habitat to
several small-sized species (López-Gappa et al., 2018; L. Schejter \&
Bremec, 2019; Sirito, 2019), and has an important role in the life
history of fishes as a food source, refuge and nursery area (Delpiani et
al., 2020; García Alonso et al., 2018; Troccoli et al., 2020). The
maintenance of this singular community is related to local and regional
oceanographic processes, including the circulation of the rich Malvinas
(Falkland) current in the area (Guerrero et al., 1999; A. R. Piola \&
Gordon, 1989) and the upwelling and mixing phenomena (Matano et al.,
2019). The input of nutrients coming from the Malvinas (Falkland)
current also supports a diverse plankton community (Guinder et al.,
2020). Overall, 811 benthic and plankton species have been identified
for the MPA N-BB ecosystem, where 349 were reported for the first time
for the area (APN, 2022). However, identifying the main species for
maintenance of ecosystem services and health as well as for management
and conservation is essential.

Recently, the structure of the southwestern South Atlantic Ocean has
been proposed to be under a `wasp-waist' control, meaning that the
structure and dynamics of the ecosystem is regulated primarily by
mid-trophic level species (e.g., fishes, crustaceans) (Padovani et al.,
2012; Riccialdelli et al., 2020; Saporiti et al., 2015). In particular,
the ecosystem of the MPA N-BB shows a more pronounced `wasp-waist'
structure than other sub-Antarctic areas, such as the Atlantic coast of
Tierra del Fuego. The Fuegian sprat \emph{Sprattus fuegensis} and
longtail southern cod \emph{Patagonotothen ramsayi} are considered the
most plausible `wasp-waist' species (Riccialdelli et al., 2020).

High-latitude marine ecosystems, such as the MPA N-BB, are characterized
by an unexpected complexity in terms of biodiversity and ecological
interactions (Cordone et al., 2020; Day et al., 2013; Kortsch et al.,
2019; Trathan et al., 2021). Although there is a robust knowledge about
the complexity considering the richness of the benthic and plankton
communities in the MPA N-BB ecosystem (APN, 2022; Guinder et al., 2020;
L. Schejter et al., 2016, 2020), there is a need for a better
understanding of species interactions' complexity and structure. This
aspect can be tackled by analyzing one of the most-frequent
relationships occurring between species: the predator-prey interaction
(Bascompte, 2009). The sum of predator-prey or trophic interactions of a
particular region is referred to as a food web, which represents the
roadmap for matter and energy flow in an ecosystem. In recent years,
network approaches have been successfully applied to study complex
high-latitude marine ecosystems, improving our knowledge on structure,
functioning, and response to environmental/anthropogenic changes
(Cordone et al., 2018; Funes et al., 2022; Kortsch et al., 2015; Marina
et al., 2023). Among anthropogenic threats, it is worth mentioning that
contaminants like mercury and microplastics have been recently reported
as important threats for the MPA N-BB region (Cossi et al., 2021; Di
Mauro et al., 2022; Fioramonti et al., 2022).

In the present work, we present the first, detailed analysis of the
network of predator-prey interactions, hereafter food web, for the MPA
N-BB ecosystem. For this, we applied a network approach to a highly
resolved food web. The objective was twofold: characterize the food web
in terms of complexity and structure, and identify the species' role in
such network framework.

@@

The MPAs Namuncurá - Banco Burdwood I and II, created by National Laws
26.875 and 27.490, comprise a shallow submarine plateau called Burdwood
Bank (BB) and a deep slope that reaches 4000 m in depth, N-BB I and N-BB
II, respectively (APN, 2023). They are located 150 km east of isla de
los Estados and 200 km south from Malvinas/Falkland Islands (Figure 1).
The MPA N-BB I comprises nearly 34,000 km\textsuperscript{2}
circumscribed by the 200 m isobath, between 54º--55ºS and 56º--62ºW,
with a slight slope extended nearly 370 km east--west. Physical features
in the BB are fairly stable, with salinity averaging 34 all year round
and temperature ranging between 4 and 8ºC (Acha et al., 2004; Guerrero
et al., 1999; A. Piola \& Falabella, 2009). The BB is surrounded by
steep flanks of up to 4000 m depth through which strong currents
circulate (Matano et al., 2019; A. R. Piola \& Gordon, 1989; Reta,
2014). Such deep slope form the basis of N-BB II that protects about
32,000 km\textsuperscript{2}. Intense upwelling and mixing occur over
it, entraining deep nutrient-rich waters into the photic layer (Matano
et al., 2019; A. Piola \& Falabella, 2009), and resulting in a fairly
homogeneous water column both spatially and temporally (Glorioso \&
Flather, 1995; Guerrero et al., 1999; Matano et al., 2019).

Given the evidence collected during several research cruises about the
oceanographic and ecological processes connecting MPAs N-BB I and II
(references in APN, 2022), a joint management plan was recently proposed
(APN, 2022).

@@

In order to build the network of predator-prey interactions we reviewed
more than 150 references considering published articles, databases and
doctoral theses. Furthermore, we took into account personal
communications from experts belonging to the working group of the study
area (https://www.pampazul.gob.ar/tag/banco-burdwood/). The diversity of
the expertise of the authors contributing to the present study was a key
factor in enhancing the quality of the network, and inherently improved
the network representation. A list of the references used to build the
network is presented in Supplementary Material (Table S1).

Due to a lack of trophic data resolution for some species inhabiting the
study area, we followed the concept of trophic species, here defined as
aggregated groups of taxa. In most cases, we followed it when specific
data on species, in the taxonomic sense, were not available. In some
cases, we collapsed species when taxa shared the same set of predators
and prey (trophic similarity, Martinez (1991)), one of the aggregation
methods that better preserves food web functional properties (Gauzens et
al., 2013). Details about this can be found in Supplementary Material
(Table S2).

With the gathered trophic data we constructed a matrix of pairwise
interactions; a value of 1 or 0 was assigned to each element a\_ij of
the matrix depending on whether the j-species preyed or not on the
i-species. Then we transformed such matrix into an oriented graph with L
trophic interactions between S nodes or species. The orientation or
direction of the graph follows the flow of energy and matter in the
network, from prey to predator.

@@

The network-level analysis aims to characterizee the food web in terms
of complexity and structure. For this we calculated several network
properties commonly used to describe empirical food webs (Pascual \&
Dunne, 2005): (1) number of species S; (2) number of interactions or
links L; (3) link density L/S; (4) connectance L/S\^{}2; (5) omnivory
Omn; and (6) small-world pattern. In order to explore the small-world
phenomenon we analyzed the characteristic path length (CPL) and the
clustering coefficient (CC). The CPL is defined as the average shortest
path length between all pairs of nodes (Watts \& Strogatz, 1998). Here,
CPL was calculated as the average number of nodes in the shortest path
\(CPL_{Min} (i,j)\) between all pairs of nodes \(S(i,j)\) in a network
averaged over \(S(S-1)/2\) nodes:

@@

\[
CC_i = \frac{2E_i}{K_i(K_i-1)}
\] where \(E_i\) is the effective number of interactions between \(K_i\)
nearest-neighbor nodes of node \(i\) and the maximal possible number of
such interactions (Newman, 2003). With the aim of testing whether the
food web presented the small-world pattern, we compared the empirical
values of CPL and CC with those resulting from 1000 randomly generated
networks with the same size (S) and number of interactions (L),
following the method proposed by Marina et al. (2018).

Also, we estimated the (7) degree distributions for the food web, for
prey and predators, and for each functional group (e.g., Amphipoda,
Ascidiacea, Bivalvia, Fish, Marine mammals, Sea birds, etc.). The prey
and predator distributions indicate the frequency of prey among
predators, and viceversa; the functional group's degree shows the
distribution of interactions within groups.

The species-level analysis aims to describe the species' role in the
food web. For this, we considered the following properties: (1)
betweenness Btw; (2) closeness Cl; (3) trophic similarity TS; (4)
topological role TR; and (5) trophic level TL (Table 1). Topological
roles refers to the fact that food webs tend to naturally organize in
non-random, modular patterns, where modules are defined as a group of
species that interact more frequently among themselves than with species
that are not members of the module (Guimerà \& Nunes Amaral, 2005).
Species can play different roles in this respect, according to the
pattern of interactions they have within their own module and/or across
modules. We computed the topological role for each species, classified
as: module hub, species with a relatively high number of interactions,
but most within its own module; module specialist, species with
relatively few interactions and most within its own module; module
connector, species with relatively few interactions mainly between
modules; and network connector, species with high connectivity between
and within modules (Guimerà \& Nunes Amaral, 2005).

We also studied the relationship between species TL and the other
species properties by performing linear regression analyses. Thus, we
considered the TL as the dependent variable and the given property
(i.e.~betweenness, closeness, trophic similarity) as the independent
variable, and obtained the coefficients (slope and intercept) for the
linear model. Models were fitted using the least squares approach. We
also explored the topological role categories with the species TL. These
species-level properties provide a general appropriate description of
species' role in empirical complex food webs (Cirtwill et al., 2018).

@@

\caption{\label{tab:table2}Network-level properties of the MPA Namuncurá - Banco Burdwood food web. CPL: Characteristic Path Length; CC: Clustering Coefficient; SW: Small-World pattern. See table 1 for definitions and ecological relevance.}
\centering
\begin{tabular}[t]{r|r|r|r|r|r|r|l}
\hline
\textbf{Species} & \textbf{Interactions} & \textbf{Density} & \textbf{Connectance} & \textbf{Omnivory} & \textbf{CPL} & \textbf{CC} & \textbf{SW}\\
\hline
379 & 1778 & 4.69 & 0.01 & 0.48 & 2.99 & 0.08 & True\\
\hline
\end{tabular}
\end{table}

The degree distribution of the food web showed an asymmetric frequency
in the number of interactions, where most of the species had a
relatively low number of interactions and few species concentrated the
majority of them (Figure 3A). The distribution of prey among predators
showed that most of the consumers fed on a low number of prey whereas
few of them had multiple preys (Figure 3B). These were the top-five
predators in number of prey: \emph{Patagonotothen guntheri}
(Notothenioid fish, 52 prey), \emph{Patagonotothen ramsayi}
(Notothenioid fish, 50 prey), \emph{Dissostichus eleginoides}
(Notothenioid fish, 30 prey), \emph{Bathyraja brachyurops}
(Chondrichthyan, 30 prey), and \emph{Bathyraja griseocauda}
(Chondrichthyan, 28 prey). Following the same distribution pattern, few
prey presented multiple predators (Figure 3C). These were the top-five
prey in number of predators: Detritus (Non-living, 153 predators), the
three categories of Diatoms considered (benthic, centric and pennate, 72
predators on average), and species of the genus \emph{Euphausia}
(Zooplankton, 46 predators). Finally, taking into account the
interactions within each functional group, again the majority of the
interactions were concentrated in few species (Figure 3D). The most
evident species were: \emph{Themisto gaudichaudii} (Amphipoda),
\emph{Zygochlamys patagonica} (Bivalvia), \emph{Aspidostoma giganteum}
(Bryozoa), \emph{Munida gregaria} (Decapoda), \emph{Patagonotothen
ramsayi} and \emph{Patagonotothen guntheri} (bentho-pelagic fish),
\emph{Psychrolutes marmoratus} (demersal fish), and species of
\emph{Euphausia} (Zooplankton). Overall, there is an evident asymmetry
in the distribution of interactions among species at different levels in
the MPA Namuncurá - Banco Burdwood food web.

@@

We also found a negative significant relationship between TL and
closeness, however less evident, meaning that low-TL species are
relatively closer to any other species in the food web (Figure 4B).
Species of genera \emph{Calanus} and \emph{Euphausia}, and species of
Brachiopoda, all of them with TL \textless{} 3, registered the highest
values of closeness (Table S3).

It is noteworthy that the highest values of betweenness were shown by
species of mid-TLs (3-4), meaning that those species participated in the
highest number of shortest paths between species (Figure 4A). The
following are the species with the highest values: \emph{Patagonotothen
ramsayi}, \emph{Dissostichus eleginoides}, \emph{Salilota australis}
(fishes), \emph{Doryteuthis gahi} (Cephalopoda), and
\emph{Patagonotothen guntheri} (Notothenioid fish) (Table S3).

>>>>

\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={The complex network of trophic interactions in a subAntarctic oceanic Marine Protected Area},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}

@@

  \title{The complex network of trophic interactions in a subAntarctic
oceanic Marine Protected Area}
    \author[]{Tomás I. Marina\textsuperscript{a}%
  %
  }
   \ead{tomasimarina@gmail.com} 
    \author[]{Irene R. Schloss\textsuperscript{a,b,c}%
  %
  }

@@

  \begin{abstract}
  Globally, the total area of the world ocean designated under marine
  protection has increased in the recent decades. The majority of these
  Marine Protected Areas (MPAs) focus on the presence of particularly
  vulnerable, keystone, or charismatic species, large numbers of endemic
  species, and/or high biodiversity across taxonomic levels. In the
  sub-Antarctic region, the level of ocean protection is mainly
  associated to oceanic islands, except for the MPAs Namuncurá - Banco
  Burdwood I and II (MPA N-BB, 53º--55ºS and 56º--62ºW, respectively),
  which are associated to a submarine plateau and its southern adjacent
  deep slope region, respectively. Here, we present the first detailed
  analysis of the network of predator-prey interactions (food web) for
  the MPA N-BB ecosystem. We applied a network approach to characterize
  the food web in terms of complexity and structure, and identifying the
  species' role in such a framework. In terms of complexity, the MPA
  N-BB food web consisted of 1778 predator-prey interactions and 379
  species, with a link density of 4.69 and a connectance of 0.01. In
  terms of structure, almost half of the consumers were omnivores
  (0.48), and the network displayed a small-world pattern. These
  findings suggest that the ecosystem might be vulnerable to external
  perturbations targeting highly connected species, although structural
  properties might provide resilience and resistance, resulting in a
  rearranged structure that preserves its original functions.
  Furthermore, we identified several species as important in terms of
  different aspects of trophic structure and functioning, and response
  to perturbations. We suggest that generalist species, mainly fishes,
  play a crucial role in the ecosystem's bentho-pelagic coupling
  process. In addition, we propose that other species besides the
  longtail southern cod \emph{Patagonotothen ramsayi} and the Fueguian
  sprat \emph{Sprattus fuegensis} should be considered as relevant
  energy transfers for the ecosystem. Finally, we argue that the
  diversity of species, including both the benthic and pelagic habitats,
  is responsible for seccuring the connectivity within the food web
  against perturbations, therefore contributing to the structure and
  stability of the ecosystem.
  \end{abstract}
    \begin{keyword}
    Food web \sep Complexity \sep Structure \sep Marine Protected
Area \sep 
    Southwest Atlantic
  \end{keyword}

@@

The evidence of benefits provided by Marine Protected Areas (MPAs) as
well as the urgent need for ocean protection have driven an
unprecedented increase in the number of MPAs worldwide in recent years
(Roberts et al., 2017; Sala et al., 2018). Globally, the total area of
the world ocean designated under marine protection adds up to
approximately 29,600,000 km\textsuperscript{2}, distributed across
nearly 18,444 MPAs and covering 8.16\% of the ocean's surface (IUCN \&
UNEP-WCMC, 2023), and therefore approaching the 10\% goal of the
Convention of Biological Diversity (CBD, 2004). Despite this progress,
recent reports have shown that actual protection has been overestimated
because it includes areas that are not yet effectively protected (only
declared) as well as areas that allow significant extractive activities
(Sala et al., 2018).

In the sub-Antarctic region, the level of ocean protection is mainly
associated to oceanic islands, such as the South Georgia and South
Sandwich, Bouvet, Prince Edward, and Macquarie islands (IUCN \&
UNEP-WCMC, 2023). Interestingly, the case of the MPAs Namuncurá - Banco
Burdwood I and II (MPA N-BB), which is the focus of this work, is unique
since these MPAs are associated to a submarine plateau and its southern
adjacent deep slope region, respectively (Falabella, 2017). In addition,
such MPAs are part of a network of protected areas in the sub-Antarctic
area (jointly with MPA Yaganes) that aims to protect and strengthen the
functioning of this southern region to maintain global ocean health.

Many of these MPAs focus on the presence of particularly vulnerable,
keystone, or charismatic species, large numbers (or proportions) of
endemic species, and/or high biodiversity across taxonomic levels (Hogg
et al., 2016). Indeed, the MPA N-BB was created to protect a sensitive
and biodiverse benthic habitat that was barely known (Falabella, 2017;
L. Schejter et al., 2016). The benthic community is featured by high
biomass of vulnerable and fragile species (mainly Porifera and Cnidaria)
that meet the characteristics of vulnerable marine ecosystems (Schejter
\& Albano, 2021), here defined as sites that present taxa with a minimum
biomass of 10 kg per 1200 m\textsuperscript{2} (Commission for the
Conservation of Antarctic Marine Living Resources, 2009). Also, it
provides habitat to several small-sized species (López-Gappa et al.,
2018; L. Schejter \& Bremec, 2019; Sirito, 2019), and has an important
role in the life history of fishes as a food source, refuge and nursery
area (Covatti Ale et al., 2022; Delpiani et al., 2020; Fischer et al.,
2022; García Alonso et al., 2018; Troccoli et al., 2020;
\textbf{Florencia2023?}). The maintenance of this singular community is
related to local and regional oceanographic processes, including the
circulation of the rich Malvinas (Falkland) current in the area
(Guerrero et al., 1999; A. R. Piola \& Gordon, 1989) and the upwelling
and mixing phenomena (Matano et al., 2019). The input of nutrients from
the Malvinas (Falkland) current also supports a diverse plankton
community (Guinder et al., 2020). Overall, 811 benthic and plankton
species have been identified for the MPA N-BB ecosystem, where 349 were
reported for the first time in the area in recent years (APN, 2022).
However, identifying the main species involved in the maintenance of
ecosystem services and health as well as for management and conservation
is essential.

Recently, the structure of the southwestern South Atlantic Ocean has
been proposed to be under a `wasp-waist' control, meaning that the
structure and dynamics of the ecosystem are regulated primarily by
mid-trophic level species (e.g., fishes, crustaceans) (Padovani et al.,
2012; Riccialdelli et al., 2020; Saporiti et al., 2015). In particular,
the ecosystem of the MPA N-BB shows a more pronounced `wasp-waist'
structure than other sub-Antarctic areas, such as the Atlantic coast of
Tierra del Fuego. The Fuegian sprat \emph{Sprattus fuegensis} and
longtail southern cod \emph{Patagonotothen ramsayi} are considered the
most plausible `wasp-waist' species (Riccialdelli et al., 2020).

High-latitude marine ecosystems, such as the MPA N-BB, are characterized
by an unexpected complexity in terms of biodiversity and ecological
interactions (Cordone et al., 2020; Day et al., 2013; Kortsch et al.,
2019; Trathan et al., 2021). Although there is a robust knowledge about
the complexity considering the richness of the benthic and plankton
communities in the MPA N-BB ecosystem (APN, 2022; Guinder et al., 2020;
L. Schejter et al., 2016, 2020), a better understanding of species
interactions' complexity and structure is needed. This aspect can be
tackled by analyzing one of the most frequent relationships between
species: the predator-prey interaction (Bascompte, 2009). The sum of
predator-prey or trophic interactions of a particular region is referred
to as a food web, representing the roadmap for matter and energy flow in
an ecosystem. In recent years, network approaches have been successfully
applied to study complex high-latitude marine ecosystems, improving our
knowledge on structure, functioning, and response to
environmental/anthropogenic changes (Cordone et al., 2018; Funes et al.,
2022; Kortsch et al., 2015; Marina et al., 2023). Among anthropogenic
threats, it is worth mentioning that contaminants like mercury and
microplastics have been recently reported as important threats to the
MPA N-BB region (Cossi et al., 2021; Di Mauro et al., 2022; Fioramonti
et al., 2022).

In the present work, we present the first detailed analysis of the
network of predator-prey interactions, hereafter food web, for the MPA
N-BB ecosystem. For this, we applied a network approach to a highly
resolved food web. The objective was twofold: characterize the food web
in terms of complexity and structure, and identify the species' role in
such a network framework.

@@

The MPAs Namuncurá - Banco Burdwood I and II, created by National Laws
26.875 and 27.490, comprise a shallow submarine plateau called Burdwood
Bank (BB) and a deep slope that reaches 4000 m in depth, N-BB I and N-BB
II, respectively (APN, 2023). They are located 150 km east of Isla de
los Estados and 200 km south of Malvinas/Falkland Islands (Figure 1).
The MPA N-BB I comprises nearly 34,000 km\textsuperscript{2}
circumscribed by the 200 m isobath, between 54º--55ºS and 56º--62ºW,
with a slight slope extended nearly 370 km east--west. Physical features
in the BB are fairly stable, with salinity averaging 34 all year round
and temperature ranging between 4 and 8ºC (Acha et al., 2004; Guerrero
et al., 1999; A. Piola \& Falabella, 2009). The BB is surrounded by
steep flanks of up to 4000 m depth through which strong currents
circulate (Matano et al., 2019; A. R. Piola \& Gordon, 1989; Reta,
2014). Such deep slope forms the basis of N-BB II, protecting about
32,000 km\textsuperscript{2}. Intense upwelling and mixing occur over
it, entraining deep nutrient-rich waters into the photic layer (Matano
et al., 2019; A. Piola \& Falabella, 2009) and resulting in a fairly
homogeneous water column both spatially and temporally (Glorioso \&
Flather, 1995; Guerrero et al., 1999; Matano et al., 2019).

Given the evidence collected during several research cruises about the
oceanographic and ecological processes connecting MPAs N-BB I and II
(references in APN, 2022), a joint management plan was recently proposed
(APN, 2022). This is why, the study area of the present work includes
both MPAs.

@@

In order to build the network of predator-prey interactions, we reviewed
more than 150 references considering published articles, Ph.D.~theses,
public databases, and reports belonging to xx research cruises conducted
in the MPAs N-BB I and II. Here it is important to note that the
sampling effort was greater in the MPA N-BB I. Furthermore, we took into
account personal communications from experts belonging to the working
group of the study area
(https://www.pampazul.gob.ar/tag/banco-burdwood/). The diversity of the
authors' expertise contributing to the present study was a key factor in
enhancing the quality of the network, and inherently improved the
network representation. A list of the references used to build the
network is presented in Supplementary Material (Table S1).

Due to a lack of trophic data resolution for some species inhabiting the
study area, we followed the concept of trophic species, here defined as
aggregated groups of taxa. In most cases, we followed it when specific
data on species, in the taxonomic sense, were not available. In some
cases, we collapsed species when taxa shared the same set of predators
and prey (trophic similarity, Martinez (1991)), one of the aggregation
methods that better preserve food web functional properties (Gauzens et
al., 2013). In addition, for endemic species (e.g.~bryozoan
\emph{Burdwoodipora paguricola}) and other species with no trophic
studies so far, we inferred their feeding interactions applying a
conservative approach that assumes that the set of prey and predators
are at some point preserved in time. In those cases we gathered
information from upper taxonomic levels (i.e.~Genus, Family, Order,
Class, Phylum) as a good proxy variable (\textbf{Morales2015?};
\textbf{Pomeranz2018?}). Details about this can be found in
Supplementary Material (Table S2).

With the gathered trophic data, we constructed a matrix of pairwise
interactions; a value of 1 or 0 was assigned to each element a\_ij of
the matrix depending on whether the j-species preyed or not on the
i-species. Then we transformed such a matrix into an oriented graph with
L trophic interactions between S nodes or species. The orientation or
direction of the graph follows the flow of energy and matter in the
network, from prey to predator.

@@

The network-level analysis aims to characterize the food web in terms of
complexity and structure. For this, we calculated several network
properties commonly used to describe empirical food webs (Pascual \&
Dunne, 2005): (1) number of species S; (2) number of interactions or
links L; (3) link density L/S; (4) connectance L/S\^{}2; (5) omnivory
Omn; and (6) small-world pattern. In order to explore the small-world
phenomenon, we analyzed the characteristic path length (CPL) and the
clustering coefficient (CC). The CPL is the average shortest path length
between all pairs of nodes (Watts \& Strogatz, 1998). Here, CPL was
calculated as the average number of nodes in the shortest path
\(CPL_{Min} (i,j)\) between all pairs of nodes \(S(i,j)\) in a network
averaged over \(S(S-1)/2\) nodes:

@@

\[
CC_i = \frac{2E_i}{K_i(K_i-1)}
\] where \(E_i\) is the effective number of interactions between \(K_i\)
nearest-neighbor nodes of node \(i\) and the maximal possible number of
such interactions (Newman, 2003). To test whether the food web presented
the small-world pattern, we compared the empirical values of CPL and CC
with those resulting from 1000 randomly generated networks with the same
size (S) and number of interactions (L), following the method proposed
by Marina et al. (2018).

Also, we estimated the (7) degree distributions for the food web, prey
and predators, and each functional group (e.g., Amphipoda, Ascidiacea,
Bivalvia, fish, marine mammals, sea birds, among others). The prey and
predator distributions indicate the frequency of prey among predators,
and viceversa; the functional group's degree shows the distribution of
interactions within groups.

The species-level analysis aims to describe the species' role in the
food web. For this, we considered the following properties: (1)
betweenness Btw; (2) closeness Cl; (3) trophic similarity TS; (4)
topological role TR; and (5) trophic level TL (Table 1). Topological
roles refer to the fact that food webs tend to naturally organize in
non-random, modular patterns, where modules are defined as a group of
species that interact more frequently among themselves than with species
that are not members of the module (Guimerà \& Nunes Amaral, 2005).
Species can play different roles in this respect, according to the
pattern of interactions within their own module and/or across modules.
We computed the topological role for each species, classified as module
hub, species with a relatively high number of interactions, but most
within its own module; module specialist, species with relatively few
interactions and most within its own module; module connector, species
with relatively few interactions mainly between modules; and network
connector, species with high connectivity between and within modules
(Guimerà \& Nunes Amaral, 2005).

We also studied the relationship between species TL and the other
species properties by performing linear regression analyses. Thus, we
considered the TL as the dependent variable and the given property
(i.e.~betweenness, closeness, trophic similarity) as the independent
variable and obtained the coefficients (slope and intercept) for the
linear model. Models were fitted using the least squares approach. We
also explored the topological role categories with the species TL. These
species-level properties provide an appropriate description of species'
role in empirical complex food webs (Cirtwill et al., 2018).

@@

\caption{\label{tab:table2}Network-level properties of the MPA Namuncurá - Banco Burdwood food web. CPL: Characteristic Path Length; CC: Clustering Coefficient; SW: Small-World pattern. See table 1 for definitions and ecological relevance.}
\centering
\begin{tabular}[t]{r|r|r|r|r|r|r|l}
\hline
\textbf{Species} & \textbf{Interactions} & \textbf{Density} & \textbf{Connectance} & \textbf{Omnivory} & \textbf{CPL} & \textbf{CC} & \textbf{SW}\\
\hline
380 & 1785 & 4.7 & 0.01 & 0.48 & 2.99 & 0.08 & True\\
\hline
\end{tabular}
\end{table}

The degree distribution of the food web showed an asymmetric frequency
in the number of interactions, where most of the species had a
relatively low number of interactions and few species concentrated most
of them (Figure 3A). The distribution of prey among predators showed
that most consumers fed on a low number of prey whereas few had multiple
preys (Figure 3B). These were the top-five predators in number of prey:
yellowfin notothen \emph{Patagonotothen guntheri} (Notothenioid fish, 52
prey), rock cod \emph{Patagonotothen ramsayi} (Notothenioid fish, 50
prey), Patagonian toothfish \emph{Dissostichus eleginoides}
(Notothenioid fish, 30 prey), broad nose skate \emph{Bathyraja
brachyurops} (Chondrichthyan, 30 prey), and graytail skate
\emph{Bathyraja griseocauda} (Chondrichthyan, 28 prey). Following the
same distribution pattern, few prey presented multiple predators (Figure
3C). These were the top-five prey in number of predators: Detritus
(Non-living, 153 predators), the three categories of Diatoms considered
(benthic, centric and pennate, 72 predators on average), and species of
the genus \emph{Euphausia} (Zooplankton, 46 predators). Finally, taking
into account the interactions within each functional group, again most
interactions were concentrated in a few species (Figure 3D). The most
evident species were: \emph{Themisto gaudichaudii} (Amphipoda),
\emph{Zygochlamys patagonica} (Bivalvia), \emph{Aspidostoma giganteum}
(Bryozoa), \emph{Munida gregaria} (Decapoda), \emph{Patagonotothen
ramsayi} and \emph{Patagonotothen guntheri} (bentho-pelagic fish),
\emph{Psychrolutes marmoratus} (demersal fish), and species of
\emph{Euphausia} (Zooplankton). Overall, there is an evident asymmetry
in the distribution of interactions among species at different levels in
the MPA N-BB food web.

@@

We also found a significant negative relationship between TL and
closeness, however less evident, meaning that low-TL species are
relatively closer to any other species in the food web (Figure 4B).
Species of genera \emph{Calanus} and \emph{Euphausia}, and species of
Brachiopoda, all with TL \textless{} 3, registered the highest closeness
values (Table S3).

Notably, species of mid-TLs (3-4) showed the highest values of
betweenness, meaning that those species participated in the highest
number of shortest paths between species (Figure 4A). The following are
the species with the highest values: \emph{Patagonotothen ramsayi},
\emph{Dissostichus eleginoides}, \emph{Salilota australis} (fishes),
\emph{Doryteuthis gahi} (Cephalopoda), and \emph{Patagonotothen
guntheri} (Notothenioid fish) (Table S3).

Considering the topological role, `module specialist' species were the
most frequent and presented a wide TL range (1 - 4.77); `module hub' was
constrained to mid-TL species (2.48 - 3.92); `module connector' from low
to mid-TLs (1 - 3.86); and `network connector', the least frequent, had
all of its species in TL = 1, except for one with TL = 3.47 (Figure 4D,
see Figure S2 for species' topological roles in a food web graph
framework). Here it is important to highlight the two latter topological
roles because they are responsible for linking modules and maintaining
the connectivity of the food web: 40 species (5 network connectors + 35
module connectors) from 20 different functional groups with a TL range =
1 - 3.86. The following were the five species with a network connector
role: Detritus, Diatoms (benthic, centric and pennate), and
\emph{Patagonotothen ramsayi}. The 35 species with a module connector
role represent these functional groups: Amphipoda, Bivalvia,
Brachiopoda, Bryozoa, Cnidaria, Cumacea, Decapoda, detritus, Diatoms,
Dinoflagellates, Echinodermata, fish (bentho-pelagic, demersal,
Chondrichthyes), Foraminifera, Polychaeta, Porifera, Pycnogonida, sea
birds, and zooplankton (see Supplementary Material Table S3 for the
identity of the species).
